Abstract Spatial Specification of Stencil Computations

### 2016-06-24 paper.tex
Remove occurance of multi-set

We outline the specification language here. Section~\ref{sec:syntax}
oultines the syntax. Section~\ref{sec:semantics} defines its semantics
via a simple multi-set interpretation over indices. Section~\ref{sec:eqs} provides an
equational theory for specifications via relation ($\equiv$) and a
theory of approximation via relation ($<:$). These equations are then
proven sound with respecto the multi-set semantics of Section~\ref{sec:semantics}.

@@

\begin{definition}[Neighbourhood index and neighbourhood subscripts]
  For an array subscript $v(\bar{e})$, we say say that $e \in \bar{e}$
  is a \emph{neighbourhood index} if it is of the form 
  $e \equiv i$, $e \equiv i + a$ or $e \equiv i - a$ where $a$ is a
  constant of integer type. That is, a neighbourhood index is a
  relative index defined as a constant translation of a base induction
  variable. The relation $\equiv$ identifies terms up-to commutativity
  of $+$ and the inverse relation of $+$ and $-$ (\eg{},
  $(-b) + i \equiv i - b$).  We classify neighbourhood indices using
  the predicate \neigh{}. An array subscript $v(\var{e})$ where every
 $e \in \bar{e}$ is a neighbourhood index is called a 
 \emph{neighbourhood subscript}. 
\end{definition}

@@

We provide a semantic model of the stencil specification language as
the basis for building correct implementations, showing the soundness
of the equational theory and approximations, and as a basis for the
inference, checking, and program synthesis algorithms.
The semantics of stencil specifications can be given a denotational
model as multisets of vectors, represent array indexing.

\begin{definition}
  A specification of maximum dimensionality $n$, is modelled by a
  multi-set of column vectors of size $n$  ($n$-dimensional vectors /
  $n$-vectors for short)
  with values drawn from $\mathbb{Z} \cup \{\infty\}$, representing
  relative indices. The $\infty$ element represents non-relative
  indices (\eg{}, constants), which are seen as infinite with respect
  to the frame of reference of a relative offset from an induction
  variable.

@@

\begin{definition}The product of two models $\otimes$ is defined:
%%
\begin{align*}
M \otimes N & =
  \{(u \bowtie v)_i \mid i \in \{0, \ldots, 2^n \}, (u, v) \in
              M \times N\}^1
\end{align*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors).  Using this, the $\otimes$ operation constructs the set
of every pairwise permutation for every pair $(u, v)$ of $n$-vectors in the
Cartesian product of $M \times N$. Note the $^1$ superscript here 
on the set-builder notation which denotes that this construction
constructs a degenerate multiset (where all the multiplicities are 1).
This is important for the function of $\otimes$ in the model (see below).

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by 
non-determinstically picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:  
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} = 
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{ccc}
0 & 1 & 2 \\
0 & 1 & 5 \\
0 & 4 & 2 \\
0 & 4 & 5 \\
3 & 1 & 2 \\
3 & 1 & 5 \\
3 & 4 & 2 \\
3 & 4 & 5 
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors 
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise pemutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_i + \neg b_{i,j} v_i
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer $i$. 
Recall that the vectors are defined over the domain
$\mathbb{Z} \cup \{\infty\}$. 
For addition and multiplication, we take 
 $\infty$ to be an absorbing element. Thus, if either
vector in a pairwise permutation contains $\infty$ in
the $j^{th}$ entry, then $j^{th}$ entry of every pairwise
permutation is $\infty$, \eg{}, 
%%
\begin{equation*}
\vtwoh{0}{1} \bowtie \vtwoh{2}{\infty} = 
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{cc}
0 & \infty \\
0 & \infty \\ 
2 & \infty \\
2 & \infty
\end{array}
\right]
\end{equation*}

@@

\begin{figure}
\input{model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Multi-set model of specifications}
\label{fig:model}
\end{figure}

>>>>

We outline the specification language here. Section~\ref{sec:syntax}
oultines the syntax. Section~\ref{sec:semantics} defines its semantics
via a simple set interpretation over indices. Section~\ref{sec:eqs} provides an
equational theory for specifications via relation ($\equiv$) and a
theory of approximation via relation ($<:$). These equations are then
proven sound with respect to the set semantics of Section~\ref{sec:semantics}.

@@

\begin{definition}[Neighbourhood index and neighbourhood subscripts]
  For an array subscript $v(\bar{e})$, we say say that $e \in \bar{e}$
  is a \emph{neighbourhood index} if it is of the form 
  $e \equiv i$, $e \equiv i + a$ or $e \equiv i - a$ where $a$ is a
  constant of integer type. That is, a neighbourhood index is a
  relative index defined as a constant translation of a base induction
  variable. The relation $\equiv$ identifies terms up-to commutativity
  of $+$ and the inverse relation of $+$ and $-$ (\eg{},
  $(-b) + i \equiv i - b$).  We classify neighbourhood indices using
  the predicate \neigh{}. An array subscript $v(\bar{e})$ where every
 $e \in \bar{e}$ is a neighbourhood index is called a 
 \emph{neighbourhood subscript}. 
\end{definition}

@@

We provide a semantic model of the stencil specification language as
the basis for building correct implementations, showing the soundness
of the equational theory and approximations, and as a basis for the
inference, checking, and program synthesis algorithms.
The semantics of stencil specifications can be given a denotational
model as sets of vectors, represent array indexing.

\begin{definition}
  A specification of maximum dimensionality $n$, is modelled by a
  set of column vectors of size $n$  ($n$-dimensional vectors /
  $n$-vectors for short)
  with values drawn from $\mathbb{Z} \cup \{\infty\}$, representing
  relative indices. The $\infty$ element represents non-relative
  indices (\eg{}, constants), which are seen as infinite with respect
  to the frame of reference of a relative offset from an induction
  variable.

@@

\begin{definition}The product of two models $\otimes$ is defined:
%%
\begin{align*}
M \otimes N & =
  \{(u \bowtie v)_i \mid i \in \{0, \ldots, 2^n \}, (u, v) \in
              M \times N\}^1
\end{align*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors).  Using this, the $\otimes$ operation constructs the set
of every pairwise permutation for every pair $(u, v)$ of $n$-vectors in the
Cartesian product of $M \times N$.

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by 
non-determinstically picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:  
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} = 
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{ccc}
0 & 1 & 2 \\
0 & 1 & 5 \\
0 & 4 & 2 \\
0 & 4 & 5 \\
3 & 1 & 2 \\
3 & 1 & 5 \\
3 & 4 & 2 \\
3 & 4 & 5 
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors 
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise pemutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_i + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer $i$. 
Recall that the vectors are defined over the domain
$\mathbb{Z} \cup \{\infty\}$. 
For addition and multiplication, we take 
 $\infty$ to be an absorbing element. Thus, if either
vector in a pairwise permutation contains $\infty$ in
the $j^{th}$ entry, then $j^{th}$ entry of every pairwise
permutation is $\infty$, \eg{}, 
%%
\begin{equation*}
\vtwoh{0}{1} \bowtie \vtwoh{2}{\infty} = 
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{cc}
0 & \infty \\
0 & \infty \\ 
2 & \infty \\
2 & \infty
\end{array}
\right]
\end{equation*}

@@

\begin{figure}
\input{model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Set model of specifications}
\label{fig:model}
\end{figure}
